Tree height and hydraulic traits shape growth responses across droughts in a temperate broadleaf forest

### 2020-03-05 manuscript/climate_sensitivity_draft.Rmd
integrated LS results comments
#80

Community-level tree growth responses to all three droughts were modest, with a mean $Rt$ of 0.86 in 1966 and 1999, and 0.84 in 1977  (Fig. 1b). In each drought, roughly 30% of the cored trees experienced $\ge$ 30% growth reductions ($Rt \le 0.7$): 29% in 1966, 32% in 1977, and 27% in 1999. However, some individuals exhibited increased growth: ($Rt > 1.0$): 26% in 1966, 22% in 1977, and 26% in 1999.

@@

Larger-diameter trees showed greater reductions in growth when evaluating the three drought years together and for 1966 individually, although DBH was not significant during 1977 or 1999 individually (Tables 1, 4). The same held true for $ln[H]$ in single-variable tests (Tables 1, 4). When combined with other predictor variables in the full models, $ln[H]$ appeared, with negative coefficient, in all full models for the three droughts combined, in the 1966 model, and in one of the two models for 1999 (Tables 1, 5).

Crown position varied as expected with height (dominant > co-dominant > intermediate > suppressed), but with substantial variation (Fig. 2d). Crown position was a much poorer predictor of $Rt$ than was height in the single-variable tests (Table 4), lending little overall support to the hypothesis that crown exposure reduces $Rt$ (Table 1). When considered alone, crown position had a significant response only in the 1966 drought, during which trees with dominant crown positiono had the lowest $Rt$. When height was included in the model, crown position was a significant predictor in the 1999 drought, with lowest $Rt$ for suppressed and then intermediate trees. Crown position was included in some of the full models (Table 5). In 1977, where height was not included in the full model, dominant trees had the lowest $Rt$, and suppressed the highest. In contrast, in full models including both height and crown position (all droughts and 1999), the lowest $Rt$ was in suppressed, followed by intermediate, trees.

@@

Resistance was negatively correlated with $ln[TWI]$ (Tables 4-5), rejecting the idea that trees in moist microsites would be less affected by drought. Nevertheless, we tested for a negative $ln[H] *ln[TWI]$ interaction, which could indicate that smaller trees (with smaller rooting volume) are more susceptible to drought in drier microenvironments with a deeper water table. This hypothesis was rejected as the $ln[H] *ln[TWI]$ interaction was never significant and had a consistently positive coefficient (Table 4).

>>>>

At the community-level, cored trees showed substantial growth reductions in all three droughts, with a mean $Rt$ of 0.86 in 1966 and 1999, and 0.84 in 1977  (Fig. 1b). In each drought, roughly 30% of the cored trees had $Rt \le 0.7$: 29% in 1966, 32% in 1977, and 27% in 1999. However, some individuals exhibited increased growth, *i.e.*, $Rt > 1.0$: 26% of trees in 1966, 22% in 1977, and 26% in 1999.

@@

Larger-diameter trees showed stronger growth reductions during drought when evaluating the three drought years together and for 1966 individually, although DBH was not significant during 1977 or 1999 individually (Tables 1, 4). The same held true for $ln[H]$ in single-variable tests (Tables 1, 4). When combined with other predictor variables in the full models, $ln[H]$ appeared, with negative coefficient, in all full models for the three droughts combined, in the 1966 model, and in one of the two models for 1999 (Tables 1, 5).

Crown position varied as expected with height (dominant > co-dominant > intermediate > suppressed), but with substantial variation (Fig. 2d). When considered alone, crown position had a significant response only in the 1966 drought, during which trees with dominant crown position had the lowest $Rt$. Crown position was a much poorer predictor of $Rt$ than was height in the single-variable tests (Table 4), lending little overall support to the hypothesis that crown exposure reduces $Rt$ (Table 1).  When height was included in the model, crown position was a significant predictor in the 1999 drought, with lowest $Rt$ for suppressed and then intermediate trees. Crown position was included in some of the full models (Table 5). In 1977, where height was not included in the full model, dominant trees had the lowest $Rt$, and suppressed trees the highest. In contrast, in full models including both height and crown position (all droughts and 1999), the lowest $Rt$ was in suppressed, followed by intermediate, trees.

@@

$Rt$ was negatively correlated with $ln[TWI]$ (Tables 4-5), rejecting the idea that trees in moist microsites would be less affected by drought. Nevertheless, we tested for a negative $ln[H] *ln[TWI]$ interaction, which could indicate that smaller trees (with smaller rooting volume) are more susceptible to drought in drier microenvironments with a deeper water table. This hypothesis was rejected as the $ln[H] *ln[TWI]$ interaction was never significant (Table 4).
